---
title: "strain_list_for_experiments"
output: html_document
---

# make a list of strains with predicted CD activity from Forster 2019 list of prevalent gut strains (from Supp. Table 5)
# strain list from Forster = "Forster_2019_SuppTable5_gutSpecies.xlsx"
# strain list that I HMMER searched = strain_list.txt
# cysteine HMMER search results = "cysteine_strains_geneHitScoreCategories_withLCD.xlsx"

# strains for cysteine experiments
```{r}
library(stringi)

# reformat strain names in strain_list and cysteine_score_filtered_wide df to replace hyphens and underscores with whitespace
# new column for species name

for (n_row in (1:nrow(cysteine_score_filtered_wide))) {
  if (grepl("-sp-", cysteine_score_filtered_wide[n_row, "strain"]) == TRUE) {
    cysteine_score_filtered_wide[n_row, "species"] <- paste(stri_split_regex(cysteine_score_filtered_wide[n_row, "strain"], "-sp-")[[1]][1], "species")

  }
  else {
    cysteine_score_filtered_wide[n_row, "species"] <- stri_replace_all_regex(cysteine_score_filtered_wide[n_row, "strain"],
                                                                 pattern = c("_"),
                                                                 replacement = c(" "))
    cysteine_score_filtered_wide[n_row, "species"] <- stri_replace_all_regex(cysteine_score_filtered_wide[n_row, "species"],
                                                                 pattern = c("-"),
                                                                 replacement = c(" "))
    cysteine_score_filtered_wide[n_row, "species"] <- sub("^(\\S*\\s+\\S+).*", "\\1", cysteine_score_filtered_wide[n_row, "species"])
  }
}

for (n_row in (1:nrow(strain_list))) {
  if (grepl("-sp-", strain_list[n_row, "X1"]) == TRUE) {
    strain_list[n_row, "species"] <- paste(stri_split_regex(strain_list[n_row, "X1"], "-sp-")[[1]][1], "species")
  }
  else {
    strain_list[n_row, "species"] <- stri_replace_all_regex(strain_list[n_row, "X1"], 
                                                                 pattern = c("_"),
                                                                 replacement = c(" "))
    strain_list[n_row, "species"] <- stri_replace_all_regex(strain_list[n_row, "species"], 
                                                                 pattern = c("-"),
                                                                 replacement = c(" "))
    strain_list[n_row, "species"] <- sub("^(\\S*\\s+\\S+).*", "\\1", strain_list[n_row, "species"])
  }
}

# excellent excellent ....

```

```{r}
# HMMER_searched_species <- strain_list$species


Forster_species <- Forster_gutSpecies$`Species Name`

Forster_gutSpecies_cysteine <- Forster_gutSpecies

cd_expected_species <- cysteine_score_filtered_wide$species

# genes <- colnames(cysteine_score_filtered_wide[,2:7])
# for (gene in genes){
#   Forster_gutSpecies[,gene] <- NA
# }

for (n_row in 1:nrow(Forster_gutSpecies_cysteine)) {
  species <- Forster_gutSpecies_cysteine[n_row, "Species Name"]
   if (species %in% cd_expected_species) {
      Forster_gutSpecies_cysteine[n_row, "predicted_cd_activity"] <- "cd expected"
      
      # for (agene in genes){
      #   if (!is.na(cysteine_score_filtered_wide[cysteine_score_filtered_wide$species == as.character(species), agene])){
      #     Forster_gutSpecies[n_row, agene] <- "pos"
      #   }
    # }
  }
}

```

# species for from propionate results
```{r}

library(stringi)

# make new species column in propionate_results

for (n_row in (1:nrow(propionate_results))) {
  if (grepl("-sp-", propionate_results[n_row, "strain"]) == TRUE) {
    propionate_results[n_row, "species"] <- paste(stri_split_regex(propionate_results[n_row, "strain"], "-sp-")[[1]][1], "species")

  }
  else {
    propionate_results[n_row, "species"] <- stri_replace_all_regex(propionate_results[n_row, "strain"],
                                                                 pattern = c("_"),
                                                                 replacement = c(" "))
    propionate_results[n_row, "species"] <- stri_replace_all_regex(propionate_results[n_row, "species"],
                                                                 pattern = c("-"),
                                                                 replacement = c(" "))
    propionate_results[n_row, "species"] <- sub("^(\\S*\\s+\\S+).*", "\\1", propionate_results[n_row, "species"])
  }
}

```

# species from Forster gut species with expected propionate production
```{r}
Forster_gutSpecies_propionate <- Forster_gutSpecies

HMMER_searched_species <- strain_list$species
Forster_species <- Forster_gutSpecies_propionate$`Species Name`
prop_expected_species <- unique(propionate_results$species)
  
Forster_gutSpecies_propionate$pathway <- NA

for (n_row in 1:nrow(Forster_gutSpecies_propionate)) {
  # check if species was HMMER searched
  species <- Forster_gutSpecies_propionate[n_row, "Species Name"]
  
  if (!species %in% HMMER_searched_species) {
    Forster_gutSpecies_propionate[n_row, "predicted_propionate_production"] <- "not searched"
  }
  else {
    if (species %in% prop_expected_species) {
      Forster_gutSpecies_propionate[n_row, "predicted_propionate_production"] <- "propionate production expected"
      
      pathway <- propionate_results$pathway[propionate_results$species == as.character(species)][1]
      Forster_gutSpecies_propionate[n_row, "pathway"] <- pathway
    }
    
    if (!species %in% prop_expected_species) {
      Forster_gutSpecies_propionate[n_row, "predicted_propionate_production"] <- "propionate production NOT expected"
    }
  }
}

```


# join cysteine and propionate species lists
```{r}

Forster_gutSpecies_cysteine_propionate <- merge(Forster_gutSpecies_cysteine, Forster_gutSpecies_propionate, by = c("Species Name", "NCBI Taxon ID"))

```

# get list of unsearched strains so I can rerun the cysteine and propionate pipelines on them
```{r}
library(readr)

Forster_gutSpecies_cysteine_propionate$strain_name <- gsub(" ", "_", Forster_gutSpecies_cysteine_propionate$`Species Name`)

unsearched_strains <- subset(Forster_gutSpecies_cysteine_propionate$strain_name, (Forster_gutSpecies_cysteine_propionate$predicted_cd_activity == "not searched"))


write_lines(unsearched_strains, file = "unsearched_strains_08082022.txt", sep = "\n")
                      
```

# merge cysteine and propionate results
```{r}
cysteine_score_filtered_wide$cysteine_expected <- "yes"
propionate_results$propionate_expected <- "yes"
cysteine_and_propionate_results <- merge(cysteine_score_filtered_wide, propionate_results, by = c("species", "strain"), all= TRUE)

cysteine_and_propionate_results_ForsterSpecies <- subset(cysteine_and_propionate_results, cysteine_and_propionate_results$species %in% Forster_gutSpecies$`Species Name`)
# View(cysteine_and_propionate_results)

write.csv(cysteine_and_propionate_results, "cysteine_and_propionate_results.csv")
write.csv(cysteine_and_propionate_results_ForsterSpecies, "cysteine_and_propionate_results_ForsterSpeices.csv")
```
